### 2014-10-09 Neutral_simulPow.Rmd
Add D1 T-test and compareTwoGrowthCurves comparisons, model description in methods

# Multifractal analysis of multispecies spatial distributions - Model simulations

I generate parameter files for the simulation of neutral/hierarchical model using logseries as the metacommunity distribution

```{r setup, eval=T }
load(".RData")
#simul  <- F # variable to perform or not the simulations

oldcd <-getwd()
source("R/Neutral_fun.r")

# Set the location of the binary 
#
neuBin <- "~/Dropbox/cpp/CaNew/Neutral/ipsNeutralExp"
neuBin64<- "~/Dropbox/cpp/CaNew/Neutral/ipsNeutralE64"

require(pander)
panderOptions('table.split.table',Inf)
options("scipen"=100, "digits"=4)
```

## Simulations with different parameters using pomac and at Time=100

Now I will simulate the model to Time=500 with a full set of parameters 
Replacement = 0,0.001,0.01,0.1,1 and ten repetitions

```{r simul256_LogSerParms, eval=F,echo=F,message=F,warning=F}
setwd("Simul")
#simul <- data.frame()
simul <- rbind(simul,simulNeutral_1Time(8,256,500))
simul <- rbind(simul,simulNeutral_1Time(64,256,500))
simul <- rbind(simul,simulNeutral_1Time(256,256,500))

setwd(oldcd)
save.image()

```

```{r simul512_LogSerParms, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

simul <- rbind(simul,simulNeutral_1Time(8,512,500))
simul <- rbind(simul,simulNeutral_1Time(64,512,500))
simul <- rbind(simul,simulNeutral_1Time(256,512,500))

# rename s/'64'/'64_512'/ *neuFish64*.txt

setwd(oldcd)
save.image()

```

## Calculate power of SAD DqSAD DqSRS for parameters

Mortality=0.2
Dispersal=0.04
Colonization=0.001

```{r power_LogSerParms, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

# Testing
#simul <- data.frame()
#sim1 <- rbind(simul,simulNeutral_1Time(65,256,100,"L",3))
#sim1 <- data.frame(nsp=65,side=256,time=100,meta="L",spMeta=87,rep=3)
#cc <- powerNeutral_1Time(simul[4,])

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

pow_AD <- data.frame()
comp_AD<- data.frame()
cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],0.2, 0.04, 0.001) # Mortality,Dispersal,Colonization
}

for(i in 1:nrow(simul)) {
pow_AD <- rbind(pow_AD, cc[[i]]$pow_AD)
comp_AD<- rbind(comp_AD,cc[[i]]$comp_AD)
}
rm(cc)
pow_AD$MrtR <- 0.2
pow_AD$DspD <- 0.04
pow_AD$ClnR <- 0.001

stopCluster(cl)

setwd(oldcd)
save.image()
```

## Calculate power of SAD DqSAD DqSRS for parameters

Mortality=0.2
Dispersal=0.4
Colonization=0.001

```{r power_LogSerParms1, eval=F,echo=F,message=T,warning=T}
setwd("Simul")

#cc <- powerNeutral_1TimeParms(simul,0.2,0.4,0.001)

#pow_AD <- rbind(pow_AD,cc$pow_AD)
#comp_AD<- rbind(comp_AD,cc$comp_AD)
#rm(cc)

mm <- 0.2
dd <- 0.4
co <- 0.001

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],mm, dd, co) # Mortality,Dispersal,Colonization
}
p_AD <- data.frame()

for(i in 1:nrow(simul)) {
	p_AD <- rbind(p_AD, cc[[i]]$pow_AD)
	comp_AD<- rbind(comp_AD,cc[[i]]$comp_AD)
}
rm(cc)
p_AD$MrtR <- mm
p_AD$DspD <- dd
p_AD$ClnR <- co

pow_AD <- rbind(pow_AD, p_AD)

stopCluster(cl)

setwd(oldcd)
save.image()
```

## Calculate power of SAD DqSAD DqSRS for parameters

Mortality=0.4
Dispersal=0.04
Colonization=0.001

```{r power_LogSerParms2, eval=F,echo=F,message=T,warning=T}
setwd("Simul")

mm <- 0.4
dd <- 0.04
co <- 0.001

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],mm, dd, co) # Mortality,Dispersal,Colonization
}
p_AD <- data.frame()

for(i in 1:nrow(simul)) {
  	p_AD <- rbind(p_AD, cc[[i]]$pow_AD)
	comp_AD<- rbind(comp_AD,cc[[i]]$comp_AD)
}
rm(cc)
p_AD$MrtR <- mm
p_AD$DspD <- dd
p_AD$ClnR <- co

pow_AD <- rbind(pow_AD, p_AD)

stopCluster(cl)

setwd(oldcd)
save.image()
```

## Calculate power of SAD DqSAD DqSRS for parameters

Mortality=0.4
Dispersal=0.4
Colonization=0.001

```{r power_LogSerParms3, eval=F,echo=F,message=T,warning=T}
setwd("Simul")

mm <- 0.4
dd <- 0.4
co <- 0.001

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],mm, dd, co) # Mortality,Dispersal,Colonization
}
p_AD <- data.frame()

for(i in 1:nrow(simul)) {
    p_AD <- rbind(p_AD, cc[[i]]$pow_AD)
	comp_AD<- rbind(comp_AD,cc[[i]]$comp_AD)
}
rm(cc)
p_AD$MrtR <- mm
p_AD$DspD <- dd
p_AD$ClnR <- co

pow_AD <- rbind(pow_AD, p_AD)

stopCluster(cl)
rm(p_AD)
setwd(oldcd)
save.image()
```
# Make plots of power!

I do not use simulations with ColonizationRate = 0.0001

Replace NumSp with spMeta=number of species in metacommunity

```{r power_PlotAD, eval=F,echo=F,message=T,warning=T}

require(ggplot2)
pow_AD$MeanSp <- as.numeric(pow_AD$MeanSp)
pow_AD$power <- as.numeric(pow_AD$power)
pow_AD$typeI <- as.numeric(pow_AD$typeI)

# Add field with combination of dispersal & mortality 
pow_AD$DspMrt <- with(pow_AD,paste0(DspD," - ",MrtR))

# Add number of species in the metacommunity
pow_AD$spMeta <- ceiling(as.numeric(pow_AD$NumSp)*1.33)

g <- ggplot(pow_AD,aes(x=spMeta,y=power)) + geom_point(shape=19,aes(size=typeI,colour=DspMrt)) + facet_grid(Side ~ Type)
g+theme_bw()

g <- ggplot(pow_AD,aes(x=MeanSp,y=typeI)) + geom_point(shape=19,aes(size=power,colour=DspMrt)) + facet_grid(Side ~ Type)
g+theme_bw()

# Recalculate power from comp_AD
#
require(plyr)
hh <-function(x) {
  t <- nrow(x)
  s <- nrow(x[x$p.value<0.05,])
  data.frame(power=s/t,n=t)
}

# Calculate power in fuction of replacement rate difference
# Falta cambiar NumSp por 
c1 <- with(comp_AD,comp_AD[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2!=RplR1,])
c1$DifR <- with(c1,abs(RplR2-RplR1))
c2 <- ddply(c1,.(NumSp,Type,DifR),hh)
#ggsave("power_DifReplacement.png", width=6,height=6,units="in",dpi=600)

g <- ggplot(c2,aes(x=DifR,y=power)) + geom_point(shape=19,position = position_jitter(height = .05),aes(colour=as.factor(Type))) + facet_grid( NumSp ~ .)
g+ scale_x_log10(breaks=c(0.001,0.01,0.09,1))+theme_bw()

# Calculate TypeI error in fuction of replacement rate
c1 <- with(comp_AD,comp_AD[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2==RplR1,])
c3 <- ddply(c1,.(NumSp,Type,RplR2),hh)
names(c3)[4]<-"typeI"
g <- ggplot(c3,aes(x=RplR2,y=typeI)) + geom_point(shape=19,aes(colour=as.factor(Type))) + facet_grid( NumSp ~ .)
g+theme_bw()

rm(c1,c2,c3)

```

DqSAD have higher power but also a very high type I error rate.
DqSRS have lower power and a lower type I error.
SAD have low power with a small number of species and a low type I error rate.

The power of the AD or KS test Depends on the number of points used so for SAD the number of species compared change the number of points used for the test, but for DqSRS and DqSAD the number of points is always q=35 thus so we could calculate more dimensions to raise the power of the test.

# Test usign glm whic factors influence power

```{r testAD_factors, eval=F,echo=F,message=T,warning=T}
require(plyr)
hh <-function(x) {
  t <- nrow(x)
  s <- nrow(x[x$p.value<0.05,])
  data.frame(count=s,power=s/t,n=t)
}
# Calculate power
c1 <- with(comp_AD,comp_AD[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2!=RplR1,])
c1$DifR <- with(c1,abs(RplR2-RplR1))
c2 <- ddply(c1,.(NumSp,Side,Type,MrtR1,DspD1,DifR),hh)
#require(dplyr)
#c2 <- tbl_df(c1)
#c3 <- c2 %>% 
#  group_by(NumSp,Side,Type,MrtR1,DspD1,DifR) %>% 
#  filter(p.value<0.05) %>% 
#  summarize(tot=n()) %>% 
#  summarize(pgt5=n()) 
model1 <-glm(count~NumSp*Side*Type*MrtR1*DspD1*DifR,poisson,data=c2)
summary(model1)
plot(model1)

# HAVE TO USE BINOMIAL !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!1

```

# Use the information dimension t-test with sd from

Fractal dimensions have several problems of estimation (Halley) so we use information dimension with sd estimated by the regression

```{r infoDim_Ttest, eval=F,echo=F,message=T,warning=T}
setwd("Simul")
#sim1 <- data.frame(nsp=65,side=256,time=100,meta="L",spMeta=87,rep=3)
# 256 = 7 points for Dq
#
#cc <- powerNeutral_1T_D1(sim1,7,1)
# 
# Dq1 <- readNeutral_calcDq("neuFish65_256T100mfOrd.txt")
# 
# D1 <- DqT[DqT$q==1,]
# dd <- D1[121:122,]
# 
# compareTwoCurvesT(dd$Type,dd$Dq,dd$SD.Dq,5)
# 
# dd <- D1[c(41,45),]
# 
# compareTwoCurvesT(dd$rep,dd$Dq,dd$SD.Dq,5)

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1T_D1(simul[i,],log2(simul$side)-1,1,0,0,0.001)
}
pow_TD1 <- data.frame()
comp_TD1<- data.frame()

for(i in 1:nrow(simul)) {
  pow_TD1 <- rbind(pow_TD1, cc[[i]]$pow_AD)
  comp_TD1<- rbind(comp_TD1,cc[[i]]$comp_AD)
}
rm(cc)

stopCluster(cl)

setwd(oldcd)
save.image()

```

The power of D1 is around .5 too low

Lets check

```{r power_PlotD1T, eval=F,echo=F,message=T,warning=T}

require(ggplot2)
pow_AD$MeanSp <- as.numeric(pow_AD$MeanSp)
pow_AD$power <- as.numeric(pow_AD$power)
pow_AD$typeI <- as.numeric(pow_AD$typeI)

# Add field with combination of dispersal & mortality 
pow_AD$DspMrt <- with(pow_AD,paste0(DspD," - ",MrtR))

# En vez de hacer merge fijarse donde genero parametros de simulacion!!!!!!!!
pad <- merge(pow_AD,simul,by.x=c("Side","NumSp"),by.y=c("side","nsp") )

g <- ggplot(pad,aes(x=spMeta,y=power)) + geom_point(shape=19,aes(size=typeI,colour=DspMrt)) + facet_grid(Side ~ Type)
g+theme_bw()

g <- ggplot(pow_AD,aes(x=MeanSp,y=typeI)) + geom_point(shape=19,aes(size=power,colour=DspMrt)) + facet_grid(Side ~ Type)
g+theme_bw()

# Recalculate power from comp_AD
#
require(plyr)
hh <-function(x) {
  t <- nrow(x)
  s <- nrow(x[x$p.value<0.05,])
  data.frame(power=s/t,n=t)
}

# Calculate power in fuction of replacement rate difference
# Falta cambiar NumSp por 
c1 <- with(comp_AD,comp_AD[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2!=RplR1,])
c1$DifR <- with(c1,abs(RplR2-RplR1))
c2 <- ddply(c1,.(NumSp,Type,DifR),hh)
#ggsave("power_DifReplacement.png", width=6,height=6,units="in",dpi=600)

g <- ggplot(c2,aes(x=DifR,y=power)) + geom_point(shape=19,position = position_jitter(height = .05),aes(colour=as.factor(Type))) + facet_grid( NumSp ~ .)
g+ scale_x_log10(breaks=c(0.001,0.01,0.09,1))+theme_bw()

# Calculate TypeI error in fuction of replacement rate
c1 <- with(comp_AD,comp_AD[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2==RplR1,])
c3 <- ddply(c1,.(NumSp,Type,RplR2),hh)
names(c3)[4]<-"typeI"
g <- ggplot(c3,aes(x=RplR2,y=typeI)) + geom_point(shape=19,aes(colour=as.factor(Type))) + facet_grid( NumSp ~ .)
g+theme_bw()

rm(c1,c2,c3)

```

```{r power_GCtest, eval=F,echo=F,message=T,warning=T}
setwd("Simul")
# sim1 <- data.frame(nsp=65,side=256,time=100,meta="L",spMeta=87,rep=3)
#
#cc <- powerNeutral_1T_GC(simul[1,],3,0,0,0.001)

# Error en
# [1] "0.2_0.04_0.0001_1_1 4 0.2_0.04_0.0001_0.1_0 3"

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1T_GC(simul[i,],3,0,0,0.001)
}
pow_GC1 <- data.frame()
comp_GC1<- data.frame()

for(i in 1:nrow(simul)) {

  pow_GC1 <- rbind(pow_GC1, cc[[i]]$pow_AD)
  comp_GC1<- rbind(comp_GC1,cc[[i]]$comp_AD)
  }

setwd(oldcd)
save.image()
```

More simulations to do comparisons with GC

```{r more_simul512_LogSerParms, eval=T,echo=F,message=F,warning=F}
setwd("Simul")

simul <- rbind(simul,simulNeutral_1Time(8,512,500,"L",20,F))
simul <- rbind(simul,simulNeutral_1Time(64,512,500,"L",20,F))
simul <- rbind(simul,simulNeutral_1Time(256,512,500,"L",20,F))

# rename s/'64'/'64_512'/ *neuFish64*.txt

setwd(oldcd)
save.image()