### 2014-10-16 Neutral_simulPow.Rmd
Change powerNeutral_1T_D1 for compute power D_1SRS & D_0SAD add more methods

# Multifractal analysis of multispecies spatial distributions - Model simulations

I generate parameter files for the simulation of neutral/hierarchical model using logseries as the metacommunity distribution

```{r setup, eval=T }
load(".RData")
#simul  <- F # variable to perform or not the simulations

oldcd <-getwd()
source("R/Neutral_fun.r")

# Set the location of the binary 
#
neuBin <- "~/Dropbox/cpp/CaNew/Neutral/ipsNeutralExp"
neuBin64<- "~/Dropbox/cpp/CaNew/Neutral/ipsNeutralE64"

require(pander)
panderOptions('table.split.table',Inf)
options("scipen"=100, "digits"=4)
```

## Neutral Simulations

Now I will simulate the model to Time=500 with a full set of parameters 
Replacement = 0,0.001,0.01,0.1,1 and ten repetitions

```{r simul256_LogSerParms, eval=F,echo=F,message=F,warning=F}
setwd("Simul")
#simul <- data.frame()
simul <- rbind(simul,simulNeutral_1Time(8,256,500))
simul <- rbind(simul,simulNeutral_1Time(64,256,500))
simul <- rbind(simul,simulNeutral_1Time(256,256,500))

setwd(oldcd)
save.image()

```

```{r simul512_LogSerParms, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

simul <- rbind(simul,simulNeutral_1Time(8,512,500))
simul <- rbind(simul,simulNeutral_1Time(64,512,500))
simul <- rbind(simul,simulNeutral_1Time(256,512,500))

# rename s/'64'/'64_512'/ *neuFish64*.txt

setwd(oldcd)
save.image()

```

## Calculate power of SAD DqSAD DqSRS using q=[-24,24]

Mortality=0.2
Dispersal=0.04
Colonization=0.001

```{r power_LogSerParms, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

# Testing
#simul <- data.frame()
#sim1 <- rbind(simul,simulNeutral_1Time(65,256,100,"L",3))
#sim1 <- data.frame(nsp=65,side=256,time=100,meta="L",spMeta=87,rep=3)
#cc <- powerNeutral_1Time(simul[4,])

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

pow_ADq24 <- data.frame()
comp_ADq24<- data.frame()
cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],0.2, 0.04, 0.001) # Mortality,Dispersal,Colonization
}

for(i in 1:nrow(simul)) {
pow_ADq24 <- rbind(pow_ADq24, cc[[i]]$pow_AD)
comp_ADq24<- rbind(comp_ADq24,cc[[i]]$comp_AD)
}
rm(cc)
pow_ADq24$MrtR <- 0.2
pow_ADq24$DspD <- 0.04
pow_ADq24$ClnR <- 0.001

stopCluster(cl)

setwd(oldcd)
save.image()
```

## Calculate power of SAD DqSAD DqSRS using q=[-24,24]

Mortality=0.2
Dispersal=0.4
Colonization=0.001

```{r power_LogSerParms1, eval=F,echo=F,message=T,warning=T}
setwd("Simul")

#cc <- powerNeutral_1TimeParms(simul,0.2,0.4,0.001)

mm <- 0.2
dd <- 0.4
co <- 0.001

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],mm, dd, co) # Mortality,Dispersal,Colonization
}
p_AD <- data.frame()

for(i in 1:nrow(simul)) {
	p_AD <- rbind(p_AD, cc[[i]]$pow_AD)
	comp_ADq24<- rbind(comp_ADq24,cc[[i]]$comp_AD)
}
rm(cc)
p_AD$MrtR <- mm
p_AD$DspD <- dd
p_AD$ClnR <- co

pow_ADq24 <- rbind(pow_ADq24, p_AD)

stopCluster(cl)

setwd(oldcd)
save.image()
```

## Calculate power of SAD DqSAD DqSRS using q=[-24,24]

Mortality=0.4
Dispersal=0.04
Colonization=0.001

```{r power_LogSerParms2, eval=F,echo=F,message=T,warning=T}
setwd("Simul")

mm <- 0.4
dd <- 0.04
co <- 0.001

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],mm, dd, co) # Mortality,Dispersal,Colonization
}
p_AD <- data.frame()

for(i in 1:nrow(simul)) {
  	p_AD <- rbind(p_AD, cc[[i]]$pow_AD)
	comp_ADq24<- rbind(comp_ADq24,cc[[i]]$comp_AD)
}
rm(cc)
p_AD$MrtR <- mm
p_AD$DspD <- dd
p_AD$ClnR <- co

pow_ADq24 <- rbind(pow_ADq24, p_AD)

stopCluster(cl)

setwd(oldcd)
save.image()
```

## Calculate power of SAD DqSAD DqSRS using q=[-24,24]

Mortality=0.4
Dispersal=0.4
Colonization=0.001

```{r power_LogSerParms3, eval=F,echo=F,message=T,warning=T}
setwd("Simul")

mm <- 0.4
dd <- 0.4
co <- 0.001

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],mm, dd, co) # Mortality,Dispersal,Colonization
}
p_AD <- data.frame()

for(i in 1:nrow(simul)) {
    p_AD <- rbind(p_AD, cc[[i]]$pow_AD)
	comp_ADq24<- rbind(comp_ADq24,cc[[i]]$comp_AD)
}
rm(cc)
p_AD$MrtR <- mm
p_AD$DspD <- dd
p_AD$ClnR <- co

pow_ADq24 <- rbind(pow_ADq24, p_AD)

stopCluster(cl)
rm(p_AD)
setwd(oldcd)
save.image()
```
# Make plots of power! using q=[-24,24]

I do not use simulations with ColonizationRate = 0.0001

Replace NumSp with spMeta=number of species in metacommunity

```{r power_PlotAD, eval=F,echo=F,message=T,warning=T}

require(ggplot2)
pow_ADq24$MeanSp <- as.numeric(pow_ADq24$MeanSp)
pow_ADq24$power <- as.numeric(pow_ADq24$power)
pow_ADq24$typeI <- as.numeric(pow_ADq24$typeI)

# Add field with combination of dispersal & mortality 
pow_ADq24$DspMrt <- with(pow_ADq24,paste0(DspD," - ",MrtR))

# Add number of species in the metacommunity
pow_ADq24$spMeta <- ceiling(as.numeric(pow_ADq24$NumSp)*1.33)

g <- ggplot(pow_ADq24,aes(x=spMeta,y=power)) + geom_point(shape=19,aes(size=typeI,colour=DspMrt)) + facet_grid(Side ~ Type) +
  ylab(bquote("Rejection Rate of"~H[0]~"(" ~alpha~"= 0.05)")) +
  xlab("Metacommunity species number") +
  scale_size_continuous(name="Type I error") +
  scale_colour_discrete(name="Disp - Mort")

g+theme_bw()
ggsave("powAD_spMeta_Side.png", width=6,height=6,units="in",dpi=600)

g <- ggplot(pow_ADq24,aes(x=MeanSp,y=typeI)) + geom_point(shape=19,aes(size=power,colour=DspMrt)) + facet_grid(Side ~ Type)
g+theme_bw()

# Recalculate power from comp_ADq24
#
require(plyr)
hh <-function(x) {
  t <- nrow(x)
  s <- nrow(x[x$p.value<0.05,])
  data.frame(power=s/t,n=t)
}

comp_ADq24$spMeta <- ceiling(as.numeric(comp_ADq24$NumSp)*1.33)

# Calculate power in fuction of replacement rate difference
# Falta cambiar NumSp por 
c1 <- with(comp_ADq24,comp_ADq24[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2!=RplR1,])
c1$DifR <- with(c1,abs(RplR2-RplR1))
c2 <- ddply(c1,.(NumSp,Type,DifR),hh)

g <- ggplot(c2,aes(x=DifR,y=power)) + geom_point(shape=19,position = position_jitter(height = .05),aes(colour=as.factor(Type))) + facet_grid( NumSp ~ .)+
  ylab(bquote("Rejection Rate of"~H[0]~"(" ~alpha~"= 0.05)")) +
  xlab(bquote(Delta ~"Replacement rate")) +
  scale_size_continuous(name="Type I error") +
  scale_colour_discrete(name="")

g+ scale_x_log10(breaks=c(0.001,0.01,0.09,1))+theme_bw()

ggsave("powAD_DifReplacement.png", width=6,height=6,units="in",dpi=600)

require(pander)
pandoc.table(c2,style="grid")

# Calculate TypeI error in fuction of replacement rate
c1 <- with(comp_ADq24,comp_ADq24[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2==RplR1,])
c3 <- ddply(c1,.(NumSp,Type,RplR2),hh)
names(c3)[4]<-"typeI"
g <- ggplot(c3,aes(x=RplR2,y=typeI)) + geom_point(shape=19,aes(colour=as.factor(Type))) + facet_grid( NumSp ~ .)
g+theme_bw()

rm(c1,c2,c3)

```

DqSAD have higher power but also a very high type I error rate.
DqSRS have lower power and a lower type I error.
SAD have low power with a small number of species and a low type I error rate.

The power of the AD or KS test Depends on the number of points used so for SAD the number of species compared change the number of points used for the test, but for DqSRS and DqSAD the number of points is always q=35 thus so we could calculate more dimensions to raise the power of the test.

# Test usign glm which factors influence power

```{r testAD_factors, eval=F,echo=F,message=T,warning=T}
require(plyr)
hh <-function(x) {
  t <- nrow(x)
  s <- nrow(x[x$p.value<0.05,])
  data.frame(count=s,power=s/t,n=t)
}
# Calculate power
c1 <- with(comp_ADq24,comp_ADq24[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2!=RplR1,])
c1$DifR <- with(c1,abs(RplR2-RplR1))
c2 <- ddply(c1,.(NumSp,Side,Type,MrtR1,DspD1,DifR),hh)
#require(dplyr)
#c2 <- tbl_df(c1)
#c3 <- c2 %>% 
#  group_by(NumSp,Side,Type,MrtR1,DspD1,DifR) %>% 
#  filter(p.value<0.05) %>% 
#  summarize(tot=n()) %>% 
#  summarize(pgt5=n()) 
model1 <-glm(count~NumSp*Side*Type*MrtR1*DspD1*DifR,poisson,data=c2)
summary(model1)
plot(model1)

# HAVE TO USE BINOMIAL !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!1

```

# Use the information dimension t-test

Fractal dimensions have several problems of estimation (Halley) so we use information dimension with sd estimated by the regression

```{r infoDim_Ttest, eval=F,echo=F,message=T,warning=T}
setwd("Simul")
#sim1 <- data.frame(nsp=65,side=256,time=100,meta="L",spMeta=87,rep=3)
# 256 = 7 points for Dq
#
#cc <- powerNeutral_1T_D1(sim1,7,1)

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1T_D1(simul[i,],log2(simul$side)-1,1,0.2,0.04,0.001)
}
pow_TD1 <- data.frame()
comp_TD1<- data.frame()

for(i in 1:nrow(simul)) {
  pow_TD1 <- rbind(pow_TD1, cc[[i]]$pow_AD)
  comp_TD1<- rbind(comp_TD1,cc[[i]]$comp_AD)
}
rm(cc)

stopCluster(cl)

setwd(oldcd)
save.image()

```

## Plot power of D1

```{r power_PlotD1T, eval=F,echo=F,message=T,warning=T}

require(ggplot2)
pow_TD1$MeanSp <- as.numeric(pow_TD1$MeanSp)
pow_TD1$power <- as.numeric(pow_TD1$power)
pow_TD1$typeI <- as.numeric(pow_TD1$typeI)
# Add number of species in the metacommunity
pow_TD1$spMeta <- ceiling(as.numeric(pow_TD1$NumSp)*1.33)

g <- ggplot(pow_TD1,aes(x=spMeta,y=power)) + geom_point(shape=19,aes(size=typeI,colour=Type)) + facet_grid(Side ~ . ) +
  ylab(bquote("Rejection Rate of"~H[0]~"(" ~alpha~"= 0.05)")) +
  xlab("Metacommunity species number") +
  scale_size_continuous(name="Type I error") +
  scale_colour_discrete(name="")

g+theme_bw()

ggsave("powTD1_spMeta_Side.png", width=6,height=6,units="in",dpi=600)

g <- ggplot(pow_TD1,aes(x=spMeta,y=typeI)) + geom_point(shape=19,aes(size=power)) + facet_grid(Side ~ Type)
g+theme_bw()

# Recalculate power from comp_AD
#
require(plyr)
hh <-function(x) {
  t <- nrow(x)
  s <- nrow(x[x$p.value<0.05,])
  data.frame(power=s/t,n=t)
}

# Calculate power in fuction of replacement rate difference
# Falta cambiar NumSp por 
comp_TD1$spMeta <- ceiling(as.numeric(comp_TD1$NumSp)*1.33)

c1 <- with(comp_TD1,comp_TD1[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2!=RplR1,])
c1$DifR <- with(c1,abs(RplR2-RplR1))

c2 <- ddply(c1,.(spMeta,Type,DifR),hh)

g <- ggplot(c2,aes(x=DifR,y=power)) + geom_point(shape=19,position = position_jitter(height = .05),aes(colour=as.factor(Type))) + facet_grid( spMeta ~ .) +
  ylab(bquote("Rejection Rate of"~H[0]~"(" ~alpha~"= 0.05)")) +
  xlab(bquote(Delta ~"Replacement rate")) +
  scale_size_continuous(name="Type I error") +
  scale_colour_discrete(name="")

g+ scale_x_log10(breaks=c(0.001,0.01,0.09,1))+theme_bw()

ggsave("powTD1_DifReplacement.png", width=6,height=6,units="in",dpi=600)

# Calculate TypeI error in fuction of replacement rate
c1 <- with(comp_TD1,comp_TD1[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2==RplR1,])
c3 <- ddply(c1,.(NumSp,Type,RplR2),hh)
names(c3)[4]<-"typeI"
g <- ggplot(c3,aes(x=RplR2,y=typeI)) + geom_point(shape=19,aes(colour=as.factor(Type))) + facet_grid( NumSp ~ .)
g+theme_bw()

rm(c1,c2,c3)

```

The power of D1 in combination with a T-test is relatively low with a maximun of 0.6 for

```{r power_GCtest, eval=F,echo=F,message=T,warning=T}
setwd("Simul")
# sim1 <- data.frame(nsp=65,side=256,time=100,meta="L",spMeta=87,rep=3)
#
#cc <- powerNeutral_1T_GC(simul[1,],3,0,0,0.001)

# Error en
# [1] "0.2_0.04_0.0001_1_1 4 0.2_0.04_0.0001_0.1_0 3"

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1T_GC(simul[i,],3,0,0,0.001)
}
pow_GC1 <- data.frame()
comp_GC1<- data.frame()

for(i in 1:nrow(simul)) {

  pow_GC1 <- rbind(pow_GC1, cc[[i]]$pow_AD)
  comp_GC1<- rbind(comp_GC1,cc[[i]]$comp_AD)
  }

stopCluster(cl)

setwd(oldcd)
save.image()
```

```{r power_PlotGC, eval=F,echo=F,message=T,warning=T}

require(ggplot2)
pow_GC1$MeanSp <- as.numeric(pow_GC1$MeanSp)
pow_GC1$power <- as.numeric(pow_GC1$power)
pow_GC1$typeI <- as.numeric(pow_GC1$typeI)

# Add number of species in the metacommunity
pow_GC1$spMeta <- ceiling(as.numeric(pow_GC1$NumSp)*1.33)

g <- ggplot(pow_GC1,aes(x=spMeta,y=power)) + geom_point(shape=19,aes(size=typeI,colour=Type)) + facet_grid(Side ~ . ) +
  ylab(bquote("Rejection Rate of"~H[0]~"(" ~alpha~"= 0.05)")) +
  xlab("Metacommunity species number") +
  scale_size_continuous(name="Type I error") +
  scale_colour_discrete(name="") 
g+theme_bw()
ggsave("powGC_spMeta_Side.png", width=6,height=6,units="in",dpi=600)

# Recalculate power from comp_AD
#
require(plyr)
hh <-function(x) {
  t <- nrow(x)
  s <- nrow(x[x$p.value<0.05,])
  data.frame(power=s/t,n=t)
}

# Calculate power in fuction of replacement rate difference
comp_GC1$spMeta <- ceiling(as.numeric(comp_GC1$NumSp)*1.33)

c1 <- with(comp_GC1,comp_GC1[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2!=RplR1,])
c1$DifR <- with(c1,abs(RplR2-RplR1))

c2 <- ddply(c1,.(spMeta,Type,DifR),hh)

g <- ggplot(c2,aes(x=DifR,y=power)) + geom_point(shape=19,position = position_jitter(height = .05),aes(colour=as.factor(Type))) + facet_grid( spMeta ~ .) +
  ylab(bquote("Rejection Rate of"~H[0]~"(" ~alpha~"= 0.05)")) +
  xlab(bquote(Delta ~"Replacement rate")) +
  scale_size_continuous(name="Type I error") +
  scale_colour_discrete(name="")

g+ scale_x_log10(breaks=c(0.001,0.01,0.09,1))+theme_bw()

ggsave("powGC_DifReplacement.png", width=6,height=6,units="in",dpi=600)

```

More simulations to do comparisons with GC

```{r more_simul512_LogSerParms, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

ps <- data.frame(nsp=c(8,64,256),si=rep(c(512,256),each=3))
require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)

simul <- foreach(i=1:nrow(ps),.combine='rbind') %dopar%
{
    simulNeutral_1Time(ps$nsp[i],ps$si[i],500,"L",20,F)
}
stopCluster(cl)

setwd(oldcd)
rm(ps,cn,cl)
save.image()
```

Compare using the GC with q<=10

```{r power_GCq10test, eval=T,echo=F,message=T,warning=T}
setwd("Simul")

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1T_GC(simul[i,],3,40.2,0.0,0.001,10)
}
pow_GCq10 <- data.frame()
comp_GCq10<- data.frame()

for(i in 1:nrow(simul)) {

  pow_GCq10 <- rbind(pow_GCq10, cc[[i]]$pow_AD)
  comp_GCq10<- rbind(comp_GCq10,cc[[i]]$comp_AD)
  }

stopCluster(cl)

setwd(oldcd)
rm(cc)
save.image()

```

The power of compareGrowthCurves is much much greater using the range -10 10 than -24 24, and type I error is lower!!

Lets try with ANderson-Darling test!

```{r power_ADq10test, eval=F,echo=F,message=T,warning=T}
setwd("Simul")

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],0.2,0.04,0.001,10)
}
pow_ADq10 <- data.frame()
comp_ADq10<- data.frame()

for(i in 1:nrow(simul)) {

  pow_ADq10 <- rbind(pow_ADq10, cc[[i]]$pow_AD)
  comp_ADq10<- rbind(comp_ADq10,cc[[i]]$comp_AD)
  }

stopCluster(cl)

setwd(oldcd)
rm(cc)
save.image()

```