### 2015-03-10 SP_simul.Rmd
Update reviewer 1 specific comments #7 #8

# Multifractal analysis of multispecies spatial distributions - Spatial patterns simulations

I use mfSBA for multifractal estimation: <https://github.com/lsaravia/mfsba>

I generate different a spatial patterns to compare SRSDq SADdq and SAD

```{r setup, eval=T }
load(".RData")

simul  <- F # variable to perform or not the simulations

oldcd <-getwd()
source("R/Neutral_fun.r")

# Set the location of the binary 
#
mfBin <- "~/Dropbox/cpp/SpatialAnalysis/mfsba" # /mfSBA /multiSpeciesSBA

require(pander)
panderOptions('table.split.table',Inf)
options("scipen"=100, "digits"=4)

```

# Banded pattern of species with uniform  SAD 256 side

First generate a pattern with  SAR z=1 using simply bands of species z=0.5 because I am using area as Borda-de-Agua instead of side.

```{r simul_z1, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

DqF <- compMethods_UniformSAD(8,256)$Dq

DqF <- rbind(DqF,compMethods_UniformSAD(64,256)$Dq)

DqF <- rbind(DqF,compMethods_UniformSAD(256,256)$Dq)

# Analizing this last it means that the scaling is broken at side 16
#  
#Dq2<- calcDq_multiSBA(fname1,"q.sed 16 1024 20 E",mfBin,T)
#Dq2$Type <- "rnzDqSAD16-512"
#Dq3<- rbind(Dq3,Dq2)
#
##Dq2<- calcDq_multiSBA(fname1,"q.sed 2 16 20 E",mfBin,T)
#Dq2$Type <- "rnzDqSAD2-16"
#Dq3<- rbind(Dq3,Dq2)
#plotDq(Dq3,"Type")
require(pander)
pandoc.table(DqF[DqF$R.Dq<0.8,],caption="R2<0.8")
setwd(oldcd)
```

* For SRS the positive part seems to reflect the equal SAD and the negative the spatial pattern

# Banded pattern 64 species 512 side

To see if there is any efect of side

```{r simul_z1_512, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

DqF <- rbind(DqF,compMethods_UniformSAD(8,512))

DqF <- rbind(DqF,compMethods_UniformSAD(64,512))

DqF <- rbind(DqF,compMethods_UniformSAD(256,512))

setwd(oldcd)
```

Thus DqSAD works in the range 2-8 2-16 2-32 depending on the number of species

For different number of species and different sides the same cualitative patterns are observed.

# Now simulate an image with logseries SAD with 8, 64 & 256 species

# Test logseries SAD side = 256

```{r simul_ls_256, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

DqF <- rbind(DqF,compMethods_FisherSAD(8,256))

DqF <- rbind(DqF,compMethods_FisherSAD(64,256))

DqF <- rbind(DqF,compMethods_FisherSAD(256,256))

setwd(oldcd)

```

# Test logseries SAD side=512

```{r simul_ls_512b, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

DqF <- rbind(DqF,compMethods_FisherSAD(8,512))

DqF <- rbind(DqF,compMethods_FisherSAD(64,512))

DqF <- rbind(DqF,compMethods_FisherSAD(256,512))

setwd(oldcd)
save.image()

```

# Simulate with logseries and the neutral model

```{r simul_neu_256, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

DqF <- rbind(DqF,compMethods_NeutralSAD(8,256,T)$Dq)

DqF <- rbind(DqF,compMethods_NeutralSAD(64,256,T)$Dq)

DqF <- rbind(DqF,compMethods_NeutralSAD(256,256,T)$Dq)
setwd(oldcd)
rm(nsp,side,spa)

save.image()
```

# Test if at time 500 the model is at steady state

Here I generate figures 1 and 2 for the appendix

```{r test_neutralTime, eval=T,echo=F,message=F,warning=F}
setwd("Simul")

sml <- T
simul <- data.frame()
require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)

parms <-expand.grid(nsp=c(8,64,256),side=c(256,512),disp=0.04,migr=0.001,repl=c(0,0.001,0.01,0.1,1))

simul <- foreach(i=1:nrow(parms),.combine='rbind') %dopar%
{
simul_NeutralPlotTime(parms$nsp[i],parms$side[i],parms$disp[i],
                                           parms$migr[i],parms$repl[i],sml,500,3,"N","L")
}

#cbPal <- c("#edf8fb","#b3cde3","#8c96c6","#8856a7","#810f7c")
cbPal <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

require(ggplot2)

g <- ggplot(simul,aes(x=Time,y=H,colour=as.factor(nsp),shape=as.factor(Rep))) + 
    geom_line() + geom_point(size=1)+
    facet_grid(ReplacementRate ~ side ) +
    ylab(bquote("Shannon index"))+ 
    scale_colour_manual(values=cbPal,name="Metacomm sp")+
    scale_shape_discrete(guide=F)

print(g+theme_bw())

setwd(oldcd)

ggsave("neuTime_side_sp_H.png", width=8,height=6,units="in",dpi=600)

g <- ggplot(simul,aes(x=Time,y=Richness,colour=as.factor(nsp),shape=as.factor(Rep))) + 
    geom_line() + geom_point(size=1)+
    facet_grid(ReplacementRate ~ side ) +
    ylab(bquote("Richness"))+ 
    scale_colour_manual(values=cbPal,name="Metacomm sp")+
    scale_shape_discrete(guide=F)
print(g+theme_bw())
ggsave("neuTime_side_sp_Rich.png", width=8,height=6,units="in",dpi=600)

save.image()

```

Perform Neutral with metaccomunity uniform and Uniform, 10 simulations with side 256 and 512

```{r test_SAD256U, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

cc <- comp_NeutralUniform(8,256)
DqUN   <-cc$Dq
SadUN  <-cc$SAD

cc <- comp_NeutralUniform(64,256)
DqUN   <-rbind(DqUN  ,cc$Dq)
SadUN  <-rbind(SadUN ,cc$SAD)

cc <- comp_NeutralUniform(256,256)
DqUN   <-rbind(DqUN  ,cc$Dq)
SadUN  <-rbind(SadUN ,cc$SAD)

cc <- comp_NeutralUniform(8,512)
DqUN   <-rbind(DqUN  ,cc$Dq)
SadUN  <-rbind(SadUN ,cc$SAD)

cc <- comp_NeutralUniform(64,512)
DqUN   <-rbind(DqUN  ,cc$Dq)
SadUN  <-rbind(SadUN ,cc$SAD)

cc <- comp_NeutralUniform(256,512)
DqUN   <-rbind(DqUN  ,cc$Dq)
SadUN  <-rbind(SadUN ,cc$SAD)

rm(cc)

# merge SAD
SadT <- rbind(SadGC,SadUN)
rm(SadGC,SadUN)

# merge Dq

DqT <- rbind(DqGC,DqUN)
rm(DqGC,DqUN)

setwd(oldcd)
save.image()

#compT <- rbind(compT,compUN)
#SadT <- rbind(SadT,SadUN)
#DqT  <- rbind(DqT,DqUN)

```

Perform Neutral with metacommunity logseries and Logseries, 10 simulations with side =256/512

```{r test_SAD256, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

cc <- comp_NeutralLogseries(8,256)
DqT   <-rbind(DqT,cc$Dq)
SadT  <-rbind(SadT ,cc$SAD)

cc <- comp_NeutralLogseries(64,256)
DqT   <-rbind(DqT  ,cc$Dq)
SadT  <-rbind(SadT ,cc$SAD)

cc <- comp_NeutralLogseries(256,256)
DqT   <-rbind(DqT  ,cc$Dq)
SadT  <-rbind(SadT ,cc$SAD)

cc <- comp_NeutralLogseries(8,512)
DqT   <-rbind(DqT  ,cc$Dq)
SadT  <-rbind(SadT ,cc$SAD)

cc <- comp_NeutralLogseries(64,512)
DqT   <-rbind(DqT  ,cc$Dq)
SadT  <-rbind(SadT ,cc$SAD)

cc <- comp_NeutralLogseries(256,512)
DqT   <-rbind(DqT  ,cc$Dq)
SadT  <-rbind(SadT ,cc$SAD)

# Simulations with side=512 not done
#
#compSP <- rbind(compSP,compKS_NeutralLogseries(8,512))
#compSP <- rbind(compSP,compKS_NeutralLogseries(64,512))
#compSP <- rbind(compSP,compKS_NeutralLogseries(256,512))
rm(cc)

setwd(oldcd)
save.image()
```

Now check if all simulations are done a make power calculations

```{r power_adtest1, eval=F,echo=F,message=F,warning=F}
# Check all simulations are present
require(plyr)
ddply(SadT,3:6,summarize, max(rep))
ddply(DqT,5:8,summarize, max(rep),min(rep))
powad <- data.frame()

powad <- rbind(powad,calcPower_AD(DqT,SadT,8,256))
powad <- rbind(powad,calcPower_AD(DqT,SadT,64,256))
powad <- rbind(powad,calcPower_AD(DqT,SadT,256,256))

powad <- rbind(powad,calcPower_AD(DqT,SadT,8,512))
powad <- rbind(powad,calcPower_AD(DqT,SadT,64,512))
powad <- rbind(powad,calcPower_AD(DqT,SadT,256,512))

save.image()
```

Make plots of power and type I error

```{r power_adtest, eval=F,echo=F,message=F,warning=F}

plot_DqPowerTypeI <- function(pp,tit){
  require(ggplot2)
  pp$Type2[pp$Type2=="SRS"] <- "DqSRS" 
  pp$Type2[grepl("rnz",pp$Type2)] <- "Power"
  pp$Type2[grepl("Dq",pp$Type2)] <- "Type I"

  g <-ggplot(pp, aes(factor(NumSp), power, colour = as.factor(Side))) +
    geom_point(aes(shape=as.factor(Side)),size=4) + ggtitle(tit) + theme_bw() +
    facet_grid(Type1~Type2) + scale_colour_discrete(name="Size") +
    scale_shape_discrete(name="Size") + scale_x_discrete(name="Number of Species")
  print(g)
}
#powad <- power
p1 <- with(powad, powad[grepl("^DqSAD|^SRS",Type1) & SAD1==SAD2 & SAD1=="Uniform" ,])
plot_DqPowerTypeI(p1,"Uniform")

p1 <- with(powad, powad[grepl("^DqSAD|^SRS",Type1) & SAD1==SAD2 & SAD1=="NeuUnif" ,])
plot_DqPowerTypeI(p1,"NeuUnif")

p1 <- with(powad, powad[grepl("^DqSAD|^SRS",Type1) & SAD1==SAD2 & SAD1=="Logseries" ,])
plot_DqPowerTypeI(p1,"Logseries")

p1 <- with(powad, powad[grepl("^DqSAD|^SRS",Type1) & SAD1==SAD2 & SAD1=="Neutral" ,])
plot_DqPowerTypeI(p1,"Neutral")

```

## Plots for publication uniform/Logseries SAD

```{r plots_spDq, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

# Plot of the Regular and Randomized spatial pattern for a uniform & Logseries SAD with 64 species and side=256
#
g <- plotSAD_SpatPat(64,256,"B")

setwd("..")
ggsave("spat_UniLog64_256.png", width=8,height=6,units="in",dpi=600)
setwd("Simul")

#
# Plots of Dq with uniform/Logseries SAD and 10 repetitions!
#
setwd("..")

g <- plotDq_Side_Sp(DqT,256,0,"B")
g
ggsave("Dq_UniLog_NumSp_256.png", width=8,height=8,units="in",dpi=600)

g <- plotDq_Side_Sp(DqT,512,0,"B")
g
ggsave("Dq_UniLog_NumSp_512.png", width=8,height=8,units="in",dpi=600)

#
# Plot R2 for dqSAD dqSRS for uniform & Logseries
#
g <- plotR2Dq_Side_Sp(DqT,256,64,"B") 
g
ggsave("R2Dq_UniLog_64_256.png", width=8,height=6,units="in",dpi=600)

g <- plotR2Dq_Side_Sp(DqT,0,0,"Uniform")
g
ggsave("R2Dq_Unif_NumSp_Side.png", width=8,height=6,units="in",dpi=600)

#
# plot of the Fit of the method for Dq
#

setwd("Simul")

zq <- readDq_fit(256,64,"Uniform")

g <- plotDqFitG(zq)

setwd(oldcd)
ggsave("DqFit_Unif_64_256.png", width=8,height=6,units="in",dpi=600)

g <- plotDqFitGQ(zq,c(-10,-4,-1,0,1,4,10))

setwd("Simul")
zq <- readDq_fit(256,8,"Uniform")
g <- plotDqFitG(zq)

setwd(oldcd)
ggsave("DqFit_Unif_8_256.png", width=8,height=6,units="in",dpi=600)

setwd("Simul")
zq <- readDq_fit(256,256,"Uniform")
g <- plotDqFitG(zq)

setwd(oldcd)
ggsave("DqFit_Unif_256_256.png", width=8,height=6,units="in",dpi=600)

```

## Plots for publication Logseries SAD

```{r plots_logser_spDq, eval=F,echo=F,message=F,warning=F}

g <- plotDq_Side_Sp(DqT,256,0,"Logseries")
g
ggsave("Dq_Logser_NumSp_256.png", width=8,height=6,units="in",dpi=600)

g <- plotDq_Side_Sp(DqT,512,0,"Logseries")
g
ggsave("Dq_Logser_NumSp_512.png", width=8,height=6,units="in",dpi=600)

#
# Plot R2 for dqSAD dqSRS
#
#g <- plotR2Dq_Side_Sp(DqT,256,64,"Logseries")
#g
#ggsave("R2Dq_Logser_64_256.png", width=8,height=6,units="in",dpi=600)

g <- plotR2Dq_Side_Sp(DqT,0,0,"Logseries")
g

g <- plotR2Dq_Side_Sp(DqT,0,0,"Neutral")
g

#
# plot of the Fit of the method for Dq
#

setwd("Simul")

zq <- readDq_fit(256,64,"Logseries")
g <-plotDqFitG(zq)
g
setwd(oldcd)
ggsave("DqFit_Logser_64_256.png", width=8,height=6,units="in",dpi=600)

setwd("Simul")

zq <- readDq_fit(256,8,"Logseries")
g <-plotDqFitG(zq)
g
setwd(oldcd)
ggsave("DqFit_Logser_8_256.png", width=8,height=6,units="in",dpi=600)

setwd("Simul")
zq <- readDq_fit(256,256,"Logseries")
g <-plotDqFitG(zq)
g
setwd(oldcd)
ggsave("DqFit_Logser_256_256.png", width=8,height=6,units="in",dpi=600)

```

## Generate Tables with R2 proportions

```{r R2tbl_Dq, eval=F,echo=F,message=F,warning=F}

#
# Generate Tables with R2 proportions
#
require(dplyr)
hh <-function(x,c,...){
  length(x[x>c])/length(x)
}

Dq1 <- filter(DqT, SAD=="Logseries" | SAD=="Uniform",Side==256,NumSp==8,Type=="DqSAD")

Dq1 <- filter(DqT, SAD=="Logseries" | SAD=="Uniform") %>%
  mutate(DqType=ifelse(grepl("SRS",Type),"SRS","SAD"), SpatialPattern=ifelse(grepl("rnz",Type),"Random","Regular")) %>%
  group_by(DqType,SAD,SpatialPattern,NumSp,Side) %>% 
  summarize(Freq60=hh(R.Dq,.6),Freq90=hh(R.Dq,.9))

write.table(format(Dq1,digits=2), file = "Dq1.dat", sep = "  ", row.names = F,quote = F)

#
# The same for neutral simulations with logseries metacommunity 
#

Dq1 <- filter(DqT, SAD=="Neutral" | SAD=="NeuUnif") %>%
  mutate(DqType=ifelse(grepl("SRS",Type),"SRS","SAD"), SpatialPattern=ifelse(grepl("rnz",Type),"Randomized","Unmodified")) %>%
  group_by(SAD,SpatialPattern,DqType,NumSp,Side) %>% 
  summarize(Freq60=hh(R.Dq,.6),Freq90=hh(R.Dq,.9))

names(SadT)
Sad1 <- filter(SadT, SAD=="Neutral") %>%
  group_by(Type,Side,NumSp,SAD,rep) %>% summarize(NumSpAct=n()) %>% summarize(meanNumSp=mean(NumSpAct),rangeSp=paste(range(NumSpAct),collapse="-"))

# Merge Sad1 with Dq1 for meanNumSp
unique(Sad1$Type)
unique(Dq1$DqType)
Dq1 <- inner_join(Dq1,Sad1) %>% select(SAD, meanNumSp,Side,SpatialPattern,DqType,Freq60,Freq90)
pow <- distinct(select(pow_ADq24,NumSp,spMeta))

Dq1 <- left_join(Dq1,pow)  %>% 
    select(SAD, spMeta,meanNumSp,Side,SpatialPattern,DqType,Freq60,Freq90)
names(Dq1)
write.table(format(Dq1,digits=2), file = "Dq1.dat", sep = "  ", row.names = F,quote = F)

rm(Dq1,Sad1,pow)
```