### 2015-05-06 ResponseToReviewers2.md
2nd review changes

# Reviewers comments resubmited manuscript

Associate Editor Comments to Author:

Associate Editor  
Comments to the Author:
I agree with the two reviewers that the manuscript has greatly improved. Please look at their responses for the few remaining issues concerning the content.
In addition, there are still many language issues and careless mistakes in the manuscript. This needs to be improved and I suggest to find (again) a native speaker for careful proofreading. I mention a few but did not check in detail.
p.4, l.2-3: check grammar
# 4, 26: change to 'first'
# 5, 13: 'graph'
# 6, 29: missing free space
# 7, 9: 'expresses'
# 7, 31: 'took'
# 11, 13: double 'with'
# 15, 21: did not 'make'
# 16, 1: What means: 'low species communities'? 
* communities with low richness

# 16, 2: What means: 'when species are around 100'?
* the number of species is around 100

# 16, 7: 'sizes'

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Corresponding Author

The author has improved the manuscript substantially from the previous version. All my concerns were addressed. The most crucial points are widened discussion of spectra interpretation and clarification of the use of Anderson-Darling criterion for spectra comparison. I believe that AD test is not an ideal solution (for example it cannot distinguish between spectra shifted along each other) but I’m now sure that there will be no inflated type 1 error.

# I think the work now is very close to be accepted but I found several places to be corrected. The most important are plots with scaling of log Zq vs log epsilon in the case of DqSAD (Fig. 3  and 6, left panels). Zq for q < 1 must increase wi th scale and Zq for q > 1 must decrease. Specifically Z0 is just a species richness and so the plot of its scaling is a usual SAR. But on the figures this pattern is reversed. I suspect there was some technical error. Otherwise description of algorithms must be reconsidered.

* For DqSAD equations 4 & 5, we have Zq(A) = Zq(e*e) is the sum of densities of species raised to q in a subplot with side e if q>0 when the area is bigger A'>A the sum Zq(A')>=Zq(A), for q<0 we have Zq(A')<=Zq(A) and this is what we can observe for DqSAD (and also DqSRS) in figures 3 and 6. 
For example if we take only the upper left square to calculate Zq(A):

1 1 2 1
2 1 2 3
2 1 2 3   
1 2 2 1

q=2 e=2  Zq(e*e)= 3*3 + 1
q=2 e=3  Zq(e*e)= 4*4 + 5*5
q=2 e=4  Zq(e*e)= 7*7 + 7*7 + 2*2

Thus Zq is increasing (the exact term is non-decreasing) when q>0

Other minor corrections:

# p. 3, l. 12: change “scaling of Renyi entropies” to “scaling exponents of Renyi entropies”
# p. 3, l. 25: Yakimov et al., 2014 and change reference accordingly
# p. 5, formula 3: I think it is better to specify that this is D1 explicitly (change Dq to D1)
# p. 5, l. 25: change “the abundance” to “the relative abundance”
# p. 6, l. 18: insert “relations” after “species-abundance-area”
# p. 11, l. 5: Fig. 3 shows not an example of fitted Dq but scaling of Zq fitted by linear regressions (and so the figure captions must be corrected)
* I corrected the captions of figures 3 and 6

# p. 14, l. 4: alfa = 0.05 obviously
# p. 14, l. 7: Table 3

Reviewer: 2

Comments to the Corresponding Author
I find the manuscript much improved, the writing more clear and the methods and results much easier to follow. I appreciate the effort the author has made to incorporate the referees criticisms. I make only some minor corrections.
.
Some comments for improvement:
# 1)      Page 12 – line 13 – “with with” should be replaced with “with”
# 2)      Page 6 – line 29 – “(Liebovitch 1998)and” should be replaced with “(Liebovitch 1998) and”
# 3)      the reference below is a book Chapter:
#Borda-de-Água, L., Hubbell, S.P. & He, F.et al. (2007). Scaling biodiversity under neutrality. Scaling 3 biodiversity, 347–375
#and should be more correctly:
#Borda-de-Água, L., Hubbell, S. P., & He, F. (2007). Scaling biodiversity under neutrality, Scaling biodiversity (eds. D. Storch, P.A. Marquet & J.H. Brown) pp. 347-375. Cambridge University Press, Cambridge.

#4)      Dal Bello, M., Maggi, E., Rindi, L., Capocchi, A., Fontanini, D., Sanz-Lazaro, C. & Benedetti-Cecchi, L. 27 (2014). Multifractal spatial distribution of epilithic microphytobenthos on a Mediterranean rocky shore. 28 Oikos, n/a—–n/a
#SHOULD BE:
#Dal Bello, M., Maggi, E., Rindi, L., Capocchi, A., Fontanini, D., Sanz-Lazaro, C. & Benedetti-Cecchi, L. 27 (2014). Multifractal spatial distribution of epilithic microphytobenthos on a Mediterranean rocky shore. Oikos, 124, 477–485.
#5)      In page 21 “Trends in ecology & evolution”, should be “Trends in Ecology & Evolution”
#6)      The reference
#Yakimov, B.N., Iudin, D.I., Solntsev, L.A. & Gelashvili, D.B. (2013). Multifractal analysis of neutral  community spatial structure. Journal of Theoretical Biology.

#Should be
#Yakimov, B.N., Iudin, D.I., Solntsev, L.A. & Gelashvili, D.B. (2013). Multifractal analysis of neutral  community spatial structure. Journal of Theoretical Biology, 343, 44–53.

My best regards to the author,

Paulo A. V. Borges

Azorean Biodiversity Group

cE3c - Centre for Ecology, Evolution and Environmental Changes
University of Azores

Associate Editor Comments to Author:

Associate Editor
Comments to the Author:
After reading the manuscript and the comments of the two reviewers I think that this approach has high potential but that the work needs some considerable improvements as well. in addition to the comments by the reviewers I suggest to take care of the following:
#- In the beginning, I had some difficulties to follow which different metrics were actually tested against each other in the end and what different information they give us about the patterns and how we could interpret them in ecological terms. There is some text abut this in the beginning of the result section but this is much too late (and to me also an odd place to put this info). I would suggest that this is made clearer early on and maybe it would be good to add a table that lists the indices together with their ecological meaning.

* I moved the paragraph about interpretation to the methods section, and added more explanations, but as I didn't know in advance how Dq will be, I present most interpretations as a discussion. I added a table with the methods used.