### 2015-05-10 Neutral_simulPow.Rmd

# Multifractal analysis of multispecies spatial distributions - Model simulations

I use here the following neutral model with exponential dispersion:

<https://github.com/lsaravia/Neutral>

I generate parameter files for the simulation of neutral/hierarchical model using logseries as the metacommunity distribution

```{r setup, eval=T }
load(".RData")
#simul  <- F # variable to perform or not the simulations

oldcd <-getwd()
source("R/Neutral_fun.r")

# Set the location of the binary 
#
neuBin <- "~/Dropbox/cpp/CaNew/Neutral/ipsNeutralExp"
neuBin64<- "~/Dropbox/cpp/CaNew/Neutral/ipsNeutralE64"

require(pander)
panderOptions('table.split.table',Inf)
options("scipen"=100, "digits"=4)
```
# Test if at time 500 the model is at steady state

Here I generate figures 1 and 2 for the appendix

```{r test_neutralTime, eval=T,echo=F,message=F,warning=F}
setwd("Simul")

sml <- T
simul <- data.frame()
require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

parms <-expand.grid(nsp=c(8,64,256),side=c(256,512),disp=0.04,migr=0.001,repl=c(0,0.001,0.01,0.1,1))

simul <- foreach(i=1:nrow(parms),.combine='rbind') %dopar%
{
simul_NeutralPlotTime(parms$nsp[i],parms$side[i],parms$disp[i],
                                           parms$migr[i],parms$repl[i],sml,500,3,"N","L")
}

#cbPal <- c("#edf8fb","#b3cde3","#8c96c6","#8856a7","#810f7c")
cbPal <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

require(ggplot2)

g <- ggplot(simul,aes(x=Time,y=H,colour=as.factor(nsp),shape=as.factor(Rep))) + 
    geom_line() + geom_point(size=1)+
    facet_grid(ReplacementRate ~ side ) +
    ylab(bquote("Shannon index"))+ 
    scale_colour_manual(values=cbPal,name="Metacomm sp")+
    scale_shape_discrete(guide=F)

print(g+theme_bw())

setwd(oldcd)

ggsave("neuTime_side_sp_H.png", width=8,height=6,units="in",dpi=600)

g <- ggplot(simul,aes(x=Time,y=Richness,colour=as.factor(nsp),shape=as.factor(Rep))) + 
    geom_line() + geom_point(size=1)+
    facet_grid(ReplacementRate ~ side ) +
    ylab(bquote("Richness"))+ 
    scale_colour_manual(values=cbPal,name="Metacomm sp")+
    scale_shape_discrete(guide=F)
print(g+theme_bw())
ggsave("neuTime_side_sp_Rich.png", width=8,height=6,units="in",dpi=600)

save.image()

```

## Neutral Simulations

Now I will simulate the model to Time=500 with a full set of parameters 
Replacement = 0,0.001,0.01,0.1,1 and 30 repetitions

```{r simul30_LogSerParms, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

ps <- data.frame(nsp=c(8,64,256),si=rep(c(512,256),each=3))
require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)

simul <- foreach(i=1:nrow(ps),.combine='rbind') %dopar%
{
    simulNeutral_1Time(ps$nsp[i],ps$si[i],500,"L",30,F)
}
stopCluster(cl)

setwd(oldcd)
rm(ps,cn,cl)
save.image()

```

## Calculate power of SAD DqSAD DqSRS using q=[-24,24]

Mortality=0.2
Dispersal=0.04
Colonization=0.001

```{r power_LogSerParms, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

# Testing
#
#simul <- data.frame()
#sim1 <- rbind(simul,simulNeutral_1Time(65,256,100,"L",3))
#sim1 <- data.frame(nsp=65,side=256,time=100,meta="L",spMeta=87,rep=3)
#cc <- powerNeutral_1Time(simul[4,])

mm <- 0.2
dd <- 0.04
co <- 0.001

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],mm, dd, co) # Mortality,Dispersal,Colonization
}
pow_ADq24 <- data.frame()
comp_ADq24<- data.frame()

for(i in 1:nrow(simul)) {
  pow_ADq24 <- rbind(pow_ADq24, cc[[i]]$pow_AD)
	comp_ADq24<- rbind(comp_ADq24,cc[[i]]$comp_AD)
}

rm(cc)

stopCluster(cl)

setwd(oldcd)
save.image()
```

DqSAD have higher power but also a very high type I error rate.
DqSRS have lower power and a lower type I error.
SAD have low power with a small number of species and a low type I error rate.

The power of the AD or KS test Depends on the number of points used so for SAD the number of species compared change the number of points used for the test, but for DqSRS and DqSAD the number of points is always q=35 thus so we could calculate more dimensions to raise the power of the test.

# Test usign glm which factors influence power

```{r testAD_factors, eval=F,echo=F,message=T,warning=T}
require(plyr)
hh <-function(x) {
  t <- nrow(x)
  s <- nrow(x[x$p.value<0.05,])
  data.frame(count=s,power=s/t,n=t)
}
# Calculate power
c1 <- with(comp_ADq24,comp_ADq24[MrtR1==MrtR2 & DspD1==DspD2 & ClnR1==ClnR2 & RplR2!=RplR1,])
c1$DifR <- with(c1,abs(RplR2-RplR1))
c2 <- ddply(c1,.(NumSp,Side,Type,DifR),hh)
#require(dplyr)
#c2 <- tbl_df(c1)
#c3 <- c2 %>% 
#  group_by(NumSp,Side,Type,MrtR1,DspD1,DifR) %>% 
#  filter(p.value<0.05) %>% 
#  summarize(tot=n()) %>% 
#  summarize(pgt5=n()) 
model1 <-glm(count~NumSp*Side*Type*DifR,poisson,data=c2)
summary(model1)
plot(model1)

```

# Power of the information dimension t-test

Fractal dimensions have several problems of estimation (Halley) so we use information dimension with sd estimated by the regression

```{r infoDim_Ttest, eval=F,echo=F,message=T,warning=T}
setwd("Simul")
#sim1 <- data.frame(nsp=65,side=256,time=100,meta="L",spMeta=87,rep=3)
# 256 = 7 points for Dq
#
#cc <- powerNeutral_1T_D1(sim1,7,1)

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1T_D1(simul[i,],log2(simul$side)-1,1,0.2,0.04,0.001)
}
pow_TD1 <- data.frame()
comp_TD1<- data.frame()

for(i in 1:nrow(simul)) {
  pow_TD1 <- rbind(pow_TD1, cc[[i]]$pow_AD)
  comp_TD1<- rbind(comp_TD1,cc[[i]]$comp_AD)
}
rm(cc)

stopCluster(cl)

setwd(oldcd)
save.image()

```

# (not used) Calculate Power using compareGrowthCurves and q=-24,24

```{r power_GCtest, eval=F,echo=F,message=T,warning=T}
setwd("Simul")
# sim1 <- data.frame(nsp=65,side=256,time=100,meta="L",spMeta=87,rep=3)
#
#cc <- powerNeutral_1T_GC(simul[1,],3,0,0,0.001)

# Error en
# [1] "0.2_0.04_0.0001_1_1 4 0.2_0.04_0.0001_0.1_0 3"

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1T_GC(simul[i,],3,0.2,0.04,0.001)
}
pow_GC1 <- data.frame()
comp_GC1<- data.frame()

for(i in 1:nrow(simul)) {

  pow_GC1 <- rbind(pow_GC1, cc[[i]]$pow_AD)
  comp_GC1<- rbind(comp_GC1,cc[[i]]$comp_AD)
  }

stopCluster(cl)

setwd(oldcd)
rm(cc)
save.image()
```

# More simulations to do comparisons with GC

```{r more_simul512_LogSerParms, eval=F,echo=F,message=F,warning=F}
setwd("Simul")

ps <- data.frame(nsp=c(8,64,256),si=rep(c(512,256),each=3))
require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)

simul <- foreach(i=1:nrow(ps),.combine='rbind') %dopar%
{
    simulNeutral_1Time(ps$nsp[i],ps$si[i],500,"L",20,F)
}
stopCluster(cl)

setwd(oldcd)
rm(ps,cn,cl)
save.image()
```

## Power using growthcurves with q<=10

```{r power_GCq10test, eval=F,echo=F,message=T,warning=T}
setwd("Simul")

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

#] "0.2_0.04_0.001_1_1 3 0.2_0.04_0.001_1_3 3"
# Error: valor ausente donde TRUE/FALSE es necesario

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1T_GC(simul[i,],3,0.2,0.04,0.001,10)
}
pow_GCq10 <- data.frame()
comp_GCq10<- data.frame()

for(i in 1:nrow(simul)) {

  pow_GCq10 <- rbind(pow_GCq10, cc[[i]]$pow_AD)
  comp_GCq10<- rbind(comp_GCq10,cc[[i]]$comp_AD)
  }

stopCluster(cl)

setwd(oldcd)
rm(cc)
save.image()

```

The power of compareGrowthCurves is much much greater using the range -10 10 than -24 24, and type I error is lower!!

## Power using Anderson-Darling test and q<=10

```{r power_ADq10test, eval=F,echo=F,message=T,warning=T}
setwd("Simul")

require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)
require(foreach)

cc <- foreach(i=1:nrow(simul)) %dopar%
{
    powerNeutral_1Time(simul[i,],0.2,0.04,0.001,10)
}
pow_ADq10 <- data.frame()
comp_ADq10<- data.frame()

for(i in 1:nrow(simul)) {

  pow_ADq10 <- rbind(pow_ADq10, cc[[i]]$pow_AD)
  comp_ADq10<- rbind(comp_ADq10,cc[[i]]$comp_AD)
  }

stopCluster(cl)

setwd(oldcd)
rm(cc)
save.image()

```

# Make plots of power! using q=[-24,24]

```{r power_PlotAD, eval=F,echo=F,message=T,warning=T}

require(dplyr)

plotPow_MeanSp_side(pow_ADq24)

ggsave("powADq24_MeanSp_Side.png", width=6,height=6,units="in",dpi=600)

require(pander)
pandoc.table(select(pow_ADq24,Side,spMeta,MeanSp, c(5:9)),style="multiline",split.tables=Inf)

pow_ADq24 <- pow_ADq24 %>% mutate(NumSp=as.numeric(NumSp),MeanSp=as.numeric(MeanSp),Side=as.numeric(Side))

comp_ADq24 <- comp_AD %>% filter(MrtR1==0.2 & DspD1==0.04 & ClnR1==0.001) %>% 
  left_join(unique(pow_ADq24[,1:3]),by=c("Side","NumSp"))

plotPow_MeanSp_difR(comp_ADq24)

ggsave("powADq24_DifReplacement.png", width=6,height=6,units="in",dpi=600)

```

## Plot power of D1

```{r power_PlotD1T, eval=F,echo=F,message=T,warning=T}

require(dplyr)

#pow_TD1$spMeta <- ceiling(as.numeric(pow_TD1$NumSp)*1.33)

pow_TD1 <- pow_TD1 %>% select(-MeanSp) %>% 
  mutate(spMeta=ceiling(as.numeric(NumSp)*1.33),power=as.numeric(power),typeI=as.numeric(typeI)) %>% 
  left_join(unique(pow_ADq10[,1:3]),by=c("Side","NumSp"))
require(pander)
pandoc.table(  select(pow_TD1,Side,spMeta,MeanSp,c(4:8)),style="multiline",split.tables=Inf)

plotPow_MeanSp_side(pow_TD1)

ggsave("powTD1_MeanSp_Side.png", width=6,height=6,units="in",dpi=600)

# Calculate power in fuction of replacement rate difference
# Falta cambiar NumSp por 
#comp_TD1$spMeta <- ceiling(as.numeric(comp_TD1$NumSp)*1.33)

pow_TD1 <- pow_TD1 %>% mutate(NumSp=as.numeric(NumSp),MeanSp=as.numeric(MeanSp),Side=as.numeric(Side))

comp_TD1 <- comp_TD1 %>% filter(MrtR1==0.2 & DspD1==0.04 & ClnR1==0.001) %>% 
  left_join(unique(pow_ADq10[,1:3]),by=c("Side","NumSp"))

plotPow_MeanSp_difR(comp_TD1)
ggsave("powTD1_DifReplacement.png", width=6,height=6,units="in",dpi=600)

```

The power of D1 in combination with a T-test is relatively low with a maximun of 0.6 for

## (Not used) Plot power of compareGrowthCurves method

```{r power_PlotGC, eval=F,echo=F,message=T,warning=T}

#pow_GC1 <-merge(pow_GC1,unique(pow_ADq10[,1:3]),by.x=c(1,2),by.y=c(1,2),all.y=F)[,c(1:2,10,4:9)]
#names(pow_GC1)[3]<- "MeanSp"
require(dplyr)

pow_GC1 <- pow_GC1 %>% select(-MeanSp) %>% 
  mutate(spMeta=ceiling(as.numeric(NumSp)*1.33),power=as.numeric(power),typeI=as.numeric(typeI)) %>% 
  left_join(unique(pow_ADq24[,1:3]),by=c("Side","NumSp")) %>%
  select(Side,spMeta,MeanSp,c(4:8))

require(pander)
pandoc.table(pow_GC1,style="multiline",split.tables=Inf)

plotPow_MeanSp_side(pow_GC1)

ggsave("powGC_spMeta_Side.png", width=6,height=6,units="in",dpi=600)

plotPow_MeanSp_difR(comp_GC1)

ggsave("powGC_DifReplacement.png", width=6,height=6,units="in",dpi=600)

```

## Plot power of Anderson-Darling test for q<=10

```{r power_PlotADq10, eval=F,echo=F,message=T,warning=T}

plotPow_MeanSp_side(pow_ADq10)

#ggsave("powADq10_MeanSp_Side.png", width=6,height=6,units="in",dpi=600)

require(plyr)
require(dplyr)
# DONE! in .RData
#
#pow_ADq10 <- pow_ADq10 %>% mutate(NumSp=as.numeric(NumSp),MeanSp=as.numeric(MeanSp),Side=as.numeric(Side))
#
#
#comp_ADq10 <- comp_ADq10 %>% filter(MrtR1==0.2 & DspD1==0.04 & ClnR1==0.001) %>% 
#  left_join(unique(pow_ADq10[,1:3]),by=c("Side","NumSp"))

#p1 <- plotPow_MeanSp_difR(comp_ADq10)
#ggsave("powADq10_DifR_256.png", width=6,height=6,units="in",dpi=600)

p1 <- plotPow_MeanSp_RplR(comp_ADq10)
#ggsave("powADq10_RplR_256.png", width=8,height=8,units="in",dpi=600)

p2 <- plotPow_MeanSp_RplR(comp_ADq10,512)
#ggsave("powADq10_RplR_512.png", width=6,height=6,units="in",dpi=600)

p1 <- rbind(p1,p2)
p1 <- select(p1,Side,spMeta,Type:n) %>% arrange(Side,spMeta,Type,DifR) 
write.table(format(p1,digits=2), file = "Dq1.dat", sep = "  ", row.names = F,quote = F)

rm(p1,p2)

require(pander)
pandoc.table( mutate(pow_ADq10 ,spMeta=ceiling(as.numeric(NumSp)*1.33),power=as.numeric(power),typeI=as.numeric(typeI)) %>% select(Side,spMeta,MeanSp,c(5:9))
,style="multiline",split.tables=Inf,round=4)
```

## Aditional simulations to plot spatial patterns with different replacement Rate $\rho$

```{r neutral_SpSimPlot256, eval=T,echo=F,message=T,warning=T}

setwd("Simul")

ps <- data.frame(nsp=c(8,64,256),si=rep(c(256,512),each=3))
require(doParallel)
cn <-detectCores()
cl <- makeCluster(cn)
registerDoParallel(cl)

sad64 <- foreach(i=1:nrow(ps),.combine='rbind') %dopar%
{
  simul_NeutralSAD(ps$nsp[i],ps$si[i],500)
}
setwd(oldcd)

```

## Plots for publication

```{r neutral_pubPlots, eval=F,echo=F,message=T,warning=T}

# Plot of the spatial pattern for Neutral model with Logseries SAD with 8,64,256 species and side=512/256
#
#g <- plotNeutral_SpatPat(64,256,500,"L",c(0.1,1))
g <- plotNeutral_SpatPat(64,256,500,"L",c(0,0.01,0.1,1))

setwd("..")
ggsave("spat_NeuLog64_256.png", width=8,height=6,units="in",dpi=600)

# Plot simulations for all metacommunities 
#
setwd("Simul")
plotNeutral_Dq(0,256,500,"L",NULL) # Null use previous simulations 
setwd("..")
ggsave("Dq_NeuLog_NumSp_256.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
plotNeutral_Dq(0,512,500,"L",NULL) # Null use previous simulations 
setwd("..")
ggsave("Dq_NeuLog_NumSp_512.png", width=8,height=8,units="in",dpi=600)

### Plot Average SAD!
setwd("Simul")
detach("package:dplyr", unload=TRUE)
den <- plotNeutral_SAD(0,256)
setwd("..")
ggsave("RAD_NeuLog_spMeta_256.png", width=8,height=6,units="in",dpi=600)

### Plot Neutral Dq Fit
#
setwd("Simul")
zq <-readNeutral_Dq_fit(64,256,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog64_256.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(8,256,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog8_256.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(256,256,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog256_256.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(8,512,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog8_512.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(64,512,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog64_512.png", width=8,height=8,units="in",dpi=600)

setwd("Simul")
zq <-readNeutral_Dq_fit(256,512,500,"L")
plotDqFitG(zq,0)
setwd(oldcd)
ggsave("DqFit_NeuLog256_512.png", width=8,height=8,units="in",dpi=600)

# Create table with R2 of neutral models Dq
#

setwd("Simul")
Dqq <- R2Neutral_Dq(256)
Dqq <- rbind(Dqq,R2Neutral_Dq(512))

Dqq <- arrange(Dqq,DqType,Side,spMeta,ReplacementRate,Freq60,Freq90) %>% select(DqType,Side,spMeta,meanSp,ReplacementRate,Freq60,Freq90)

setwd(oldcd)

#write.table(format(Dqq,digits=2), file = "Dq1.dat", sep = "  ", row.names = F,quote = F)

require(pander)
pandoc.table(format(Dqq,digits=2),style="multiline",split.tables=Inf)

# Table with mean number of species by replacement rate
#
Dqq <- select(Dqq,Side,spMeta,ReplacementRate,meanSp)

#write.table(format(Dqq,digits=2), file = "Dq1.dat", sep = "  ", row.names = F,quote = F)
pandoc.table(format(Dqq,digits=2),style="multiline",split.tables=Inf)

rm(Dqq,Dqq1)
```