### 2015-05-14 ResponseToReviewers2.md
Submited 2nd revision of the manuscript

# Reviewers comments resubmited manuscript

Associate Editor Comments to Author:

Associate Editor  
Comments to the Author:
I agree with the two reviewers that the manuscript has greatly improved. Please look at their responses for the few remaining issues concerning the content.
#In addition, there are still many language issues and careless mistakes in the manuscript. This needs to be improved and I suggest to find (again) a native speaker for careful proofreading. I mention a few but did not check in detail.
# p.4, l.2-3: check grammar
# 4, 26: change to 'first'
# 5, 13: 'graph'
# 6, 29: missing free space
# 7, 9: 'expresses'
# 7, 31: 'took'
# 11, 13: double 'with'
# 15, 21: did not 'make'
# 16, 1: What means: 'low species communities'? 
* I corrected as: communities with low richness

# 16, 2: What means: 'when species are around 100'?
* I corrected as: the number of species is around 100

# 16, 7: 'sizes'

Following the review of a native speaker I reworded the paragraph at page 6 line 29. Additionally I deleted a paragraph in the disccussion (page 16 between lines 14-15) because it is a repetition of the same ideas and seems to be out of place.

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Corresponding Author

The author has improved the manuscript substantially from the previous version. All my concerns were addressed. The most crucial points are widened discussion of spectra interpretation and clarification of the use of Anderson-Darling criterion for spectra comparison. I believe that AD test is not an ideal solution (for example it cannot distinguish between spectra shifted along each other) but I’m now sure that there will be no inflated type 1 error.

# I think the work now is very close to be accepted but I found several places to be corrected. The most important are plots with scaling of log Zq vs log epsilon in the case of DqSAD (Fig. 3  and 6, left panels). Zq for q < 1 must increase wi th scale and Zq for q > 1 must decrease. Specifically Z0 is just a species richness and so the plot of its scaling is a usual SAR. But on the figures this pattern is reversed. I suspect there was some technical error. Otherwise description of algorithms must be reconsidered.

* I corrected this. The error was produced because I use the following to calculate DqSAD: DqSAD=-Tau/(q-1) which is equivalent to DqSAD=Tau/(1-q) that is equation 5, as Tau=log(Zq)/log(e) the slopes were inverted. I updated figures 3, 6, and the appendix.

# Other minor corrections:

# p. 3, l. 12: change “scaling of Renyi entropies” to “scaling exponents of Renyi entropies”
# p. 3, l. 25: Yakimov et al., 2014 and change reference accordingly
# p. 5, formula 3: I think it is better to specify that this is D1 explicitly (change Dq to D1)
# p. 5, l. 25: change “the abundance” to “the relative abundance”
# p. 6, l. 18: insert “relations” after “species-abundance-area”
# p. 11, l. 5: Fig. 3 shows not an example of fitted Dq but scaling of Zq fitted by linear regressions (and so the figure captions must be corrected)
* I corrected the captions of figures 3 and 6

# p. 14, l. 4: alfa = 0.05 obviously
# p. 14, l. 7: Table 3

Reviewer: 2

Comments to the Corresponding Author
I find the manuscript much improved, the writing more clear and the methods and results much easier to follow. I appreciate the effort the author has made to incorporate the referees criticisms. I make only some minor corrections.
.
Some comments for improvement:
# 1)      Page 12 – line 13 – “with with” should be replaced with “with”
# 2)      Page 6 – line 29 – “(Liebovitch 1998)and” should be replaced with “(Liebovitch 1998) and”
# 3)      the reference below is a book Chapter:
#Borda-de-Água, L., Hubbell, S.P. & He, F.et al. (2007). Scaling biodiversity under neutrality. Scaling 3 biodiversity, 347–375
#and should be more correctly:
#Borda-de-Água, L., Hubbell, S. P., & He, F. (2007). Scaling biodiversity under neutrality, Scaling biodiversity (eds. D. Storch, P.A. Marquet & J.H. Brown) pp. 347-375. Cambridge University Press, Cambridge.

#4)      Dal Bello, M., Maggi, E., Rindi, L., Capocchi, A., Fontanini, D., Sanz-Lazaro, C. & Benedetti-Cecchi, L. 27 (2014). Multifractal spatial distribution of epilithic microphytobenthos on a Mediterranean rocky shore. 28 Oikos, n/a—–n/a
#SHOULD BE:
#Dal Bello, M., Maggi, E., Rindi, L., Capocchi, A., Fontanini, D., Sanz-Lazaro, C. & Benedetti-Cecchi, L. 27 (2014). Multifractal spatial distribution of epilithic microphytobenthos on a Mediterranean rocky shore. Oikos, 124, 477–485.
#5)      In page 21 “Trends in ecology & evolution”, should be “Trends in Ecology & Evolution”
#6)      The reference
#Yakimov, B.N., Iudin, D.I., Solntsev, L.A. & Gelashvili, D.B. (2013). Multifractal analysis of neutral  community spatial structure. Journal of Theoretical Biology.

#Should be
#Yakimov, B.N., Iudin, D.I., Solntsev, L.A. & Gelashvili, D.B. (2013). Multifractal analysis of neutral  community spatial structure. Journal of Theoretical Biology, 343, 44–53.

My best regards to the author,

Paulo A. V. Borges

Azorean Biodiversity Group